---
title: "Part 2"
author: "Davide Onorio"
date: "2026-04-13"
output:
  pdf_document:
    toc: true
    toc_depth: 3
    number_sections: true
    latex_engine: xelatex
header-includes:
  - \usepackage{booktabs}
  - \usepackage{float}
  - \usepackage{amsmath}
  - \usepackage{setspace}
  - \onehalfspacing
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo      = FALSE,
  warning   = FALSE,
  message   = FALSE,
  fig.pos   = "H",
  fig.align = "center",
  fig.width  = 8,
  fig.height = 5
)

library(tidyquant)
library(tidyverse)
library(forecast)
library(lmtest)
library(sandwich)
library(moments)
library(lubridate)
library(knitr)
library(kableExtra)
```

\newpage

# Part II

## Introduction

Can financial markets be beaten using only the information embedded in
past prices? Economists and practitioners have been asking this question since 
at least the 1960s, and it remains as relevant today as ever, particularly 
given the rise of algorithmic trading, machine learning, and more exotic
assets like cryptocurrencies that operate outside traditional market
microstructures.

The theoretical foundation for answering this question is the
\textbf{Efficient Market Hypothesis (EMH)}, formalised by Fama in 1970,
provides a theoretical basis for answering this question by stating that 
prices always fully reflect available information.
Under its weakest form -- \textit{weak-form efficiency} -- the relevant
information set consists solely of the history of past prices.  If this holds 
true, no investor can systematically profit from studying price charts or 
return patterns because all information from the past is already priced in.

Formally, weak-form efficiency is closely linked to the concept of a
\textbf{martingale}: a stochastic process in which the best forecast of
tomorrow's price, given all past prices, is simply today's price:
$$\mathbb{E}[p_{t+1} \mid p_t, p_{t-1}, \ldots] = p_t.$$
This implies that expected price changes are zero and, crucially, that
past returns carry no predictive power over future returns. The random
walk is a special case of this idea, and three progressively relaxed
versions are considered in this analysis:

\begin{itemize}
  \item \textbf{RW1}: price increments are i.i.d. -- the strictest assumption,
        implying both zero autocorrelation and identical distributions over time.
  \item \textbf{RW2}: increments are independent but not necessarily identically
        distributed, accommodating time-varying volatility (heteroscedasticity).
  \item \textbf{RW3}: increments are merely uncorrelated -- the weakest and most
        empirically relevant assumption for financial returns.
\end{itemize}

## Data - Asset Selection

The dataset consists of \textbf{16 financial assets} retrieved from
Yahoo Finance via the \texttt{tidyquant} package, covering the period
\textbf{January 2015 to December 2024}. Rather than restricting the
analysis to a standard universe of large-cap US equities -- where
efficiency is most expected and most studied -- the basket was
deliberately constructed to be heterogeneous across four dimensions:
\textit{asset class} (equities, cryptocurrencies, ETFs, commodities),
\textit{geography} (US, Latin America, South-East Asia, China),
\textit{sector} (semiconductors, biotech, clean energy, fintech, space,
quantum computing), and \textit{risk profile} (from defensive assets
like GLD and NEE to highly speculative ones like IONQ, ACHR, and the
two cryptocurrencies).

This diversity is intentional. Cryptocurrencies trade continuously
outside traditional exchange infrastructure and attract a largely
retail investor base; emerging market assets like NU, MELI, SE and
KWEB operate under different regulatory and liquidity conditions; and
speculative growth stocks vary enormously in analyst coverage and
institutional participation. If the EMH holds universally, it should
hold across all of these asset types -- not just in the most liquid
and well-arbitraged corners of the market. Conversely, finding
predictability in specific segments would itself be informative,
pointing to where efficiency is most fragile.

## ARMA Analysis and Martingale Hypothesis

### Theoretical Framework

A natural starting point for assessing return predictability is to ask
whether the history of past returns contains any linear information about
future returns. The \textbf{Autoregressive Moving Average (ARMA)} class
of models provides a flexible and parsimonious framework for capturing
such dynamics. An ARMA($p$,$q$) model expresses the current return as a
linear function of its own past values and past forecast errors:

$$r_t = c + \sum_{i=1}^{p} \phi_i r_{t-i} + \sum_{j=1}^{q} \theta_j
\varepsilon_{t-j} + \varepsilon_t, \qquad \varepsilon_t \sim
\text{WN}(0, \sigma^2)$$

where $\phi_1, \ldots, \phi_p$ are the autoregressive coefficients,
$\theta_1, \ldots, \theta_q$ are the moving average coefficients, and
$\varepsilon_t$ is a white noise error term. The two special cases are
the pure \textbf{AR($p$)} model ($q = 0$), which relates current returns
to $p$ of their own lags, and the pure \textbf{MA($q$)} model ($p = 0$),
which relates current returns to $q$ past shocks. The most restrictive
case, \textbf{ARMA(0,0)}, implies that returns are simply white noise,
no past information is relevant, and the series is inherently
unpredictable from its own history.

The connection to market efficiency is direct. Under the
\textbf{martingale hypothesis}, the conditional expectation of returns
given all past information equals the unconditional mean:
$$\mathbb{E}[r_t \mid r_{t-1}, r_{t-2}, \ldots] = \mathbb{E}[r_t]$$
This is equivalent to saying that all autoregressive and moving average
coefficients in the ARMA representation are jointly zero -- i.e. the
true model is ARMA(0,0). Any statistically significant ARMA structure
would therefore constitute evidence against the martingale and, by
extension, against weak-form efficiency.

### Model Selection: The Hyndman--Khandakar Algorithm

Rather than testing a pre-specified model, we let the data determine the
appropriate ARMA order using the \textbf{Hyndman--Khandakar algorithm}
(Hyndman and Khandakar, 2008), implemented via \texttt{auto.arima()} in
R. The algorithm searches over ARMA($p$,$q$) specifications and selects
the model that minimises the Akaike Information Criterion (AIC),
penalising complexity to avoid overfitting. Since log-returns are
stationary by construction, the differencing order is fixed at $d = 0$.
The selected model takes the form:

$$r_t = c + \phi_1 r_{t-1} + \cdots + \phi_p r_{t-p} +
\theta_1 \varepsilon_{t-1} + \cdots + \theta_q \varepsilon_{t-q} +
\varepsilon_t$$

where $\phi_1, \ldots, \phi_p$ are the autoregressive coefficients,
$\theta_1, \ldots, \theta_q$ are the moving average coefficients, and
$\varepsilon_t$ is white noise. The auto-selected model is then
benchmarked against two simpler alternatives -- AR(1) and MA(1) -- to
assess whether the added complexity is warranted.

Residual adequacy is evaluated using the \textbf{Ljung--Box test}, which
tests the joint null hypothesis that the first $K$ autocorrelations of
the residuals are all zero. A p-value above 0.05 confirms that the
residuals behave as white noise and that the model has captured all
linear dependence in the series.

### Results

```{r arma-results}
results <- data.frame(
  Ticker     = c("BTC-USD","AMD","KWEB","GLD","CCJ","NEE","MELI","CRSP",
                 "SE","SOL-USD","PLTR","RKLB","ACHR","IONQ","HOOD","NU"),
  Best_Model = c("ARMA(1,0)","ARMA(3,0)","ARMA(0,3)","ARMA(2,3)","ARMA(0,0)",
                 "ARMA(2,2)","ARMA(0,0)","ARMA(2,3)","ARMA(0,0)","ARMA(3,1)",
                 "ARMA(2,2)","ARMA(0,0)","ARMA(1,0)","ARMA(0,0)","ARMA(2,0)",
                 "ARMA(2,2)"),
  AIC_Auto   = c(-13804.67,-9534.55,-11468.38,-16610.74,-10883.40,-13897.43,
                 -10452.75,-7391.45,-6469.20,-4419.83,-3651.47,-3488.83,
                 -3162.51,-2891.94,-2882.06,-2764.72),
  AIC_AR1    = c(-13804.67,-9529.48,-11464.63,-16602.23,-10880.69,-13859.88,
                 -10450.78,-7377.95,-6467.57,-4416.18,-3634.33,-3485.59,
                 -3160.51,-2888.62,-2877.67,-2745.66),
  AIC_MA1    = c(-13804.61,-9529.01,-11464.69,-16602.27,-10880.68,-13859.41,
                 -10450.78,-7377.96,-6467.51,-4416.02,-3633.85,-3485.59,
                 -3160.43,-2888.62,-2877.68,-2745.69),
  LB_pvalue  = c(0.0548,0.8275,0.7562,0.7450,0.5317,0.8086,0.0036,0.7486,
                 0.6309,0.3465,0.0366,0.7863,0.2719,0.1779,0.9020,0.1675),
  WhiteNoise = c("Yes","Yes","Yes","Yes","Yes","Yes","No","Yes","Yes","Yes",
                 "No","Yes","Yes","Yes","Yes","Yes")
)

kable(results, booktabs = TRUE,
      caption = "\\label{tab:arma}ARMA Model Selection and Residual Diagnostics",
      col.names = c("Ticker","Best Model","AIC Auto","AIC AR(1)",
                    "AIC MA(1)","LB p-value","White Noise")) %>%
  kable_styling(latex_options = c("hold_position","striped","scale_down"),
                font_size = 9) %>%
  footnote(general = "LB p-value refers to the Ljung-Box test at lag 10.
           White Noise = Yes if p-value $>$ 0.05.",
           escape = FALSE)
```

Table \ref{tab:arma} reports the selected ARMA model for each asset,
alongside AIC comparisons with the AR(1) and MA(1) benchmarks and the
Ljung--Box test on residuals. Three broad findings emerge.

First, the auto-selected model consistently outperforms both benchmarks
on AIC across all 16 assets, confirming that the Hyndman--Khandakar
algorithm identifies genuinely better-fitting specifications rather than
simply adding noise. The AIC improvements are particularly large for
NEE ($\Delta \text{AIC} \approx 37$), GLD ($\Delta \approx 8$), and
CRSP ($\Delta \approx 13$), suggesting that these assets have
meaningful ARMA structure that a simple AR(1) or MA(1) would miss.

Second, the degree of complexity varies substantially across assets.
Four assets -- CCJ, RKLB, IONQ, SE, and MELI -- are best described
by an \textbf{ARMA(0,0)}, implying that their log-returns are already
white noise with no exploitable linear structure. At the other extreme,
GLD and CRSP both require \textbf{ARMA(2,3)} specifications, while NEE,
PLTR, and NU call for \textbf{ARMA(2,2)}. This is somewhat surprising:
GLD and NEE are typically considered highly liquid and efficiently
priced, yet their return dynamics exhibit richer autocorrelation
patterns than most of the speculative assets in the basket.

Third, and most importantly for the question of efficiency, 14 out of
16 assets pass the Ljung--Box test with p-values well above 0.05,
indicating that the selected models successfully absorb all linear
dependence in the return series and leave white noise residuals. The
two exceptions are \textbf{MELI} (p = 0.004) and \textbf{PLTR}
(p = 0.037), both of which reject the white noise hypothesis even after
model selection. For MELI this is particularly striking, since the
algorithm chose an ARMA(0,0) -- meaning that despite the absence of
identifiable linear structure, the residuals are still autocorrelated,
pointing to some nonlinear or structural predictability that a standard
ARMA framework cannot capture. For PLTR, the rejection is subtler but
consistent with its relatively complex ARMA(2,2) structure and will
reappear in the martingale test that follows.

## Martingale Hypothesis Test

### Theoretical Framework

The ARMA analysis in the previous section identifies the best linear
model for each return series, but does not directly test whether the
detected structure is statistically meaningful. This is the role of the
\textbf{martingale hypothesis test}. Under the martingale, past returns
should carry no predictive power over future returns:
$$\mathbb{E}[r_t \mid r_{t-1}, r_{t-2}, \ldots] = \mathbb{E}[r_t]$$
To test this formally, we specify two competing regression models for
each asset:
$$\text{Restricted:} \quad r_t = c + \varepsilon_t$$
$$\text{Unrestricted:} \quad r_t = c + \phi_1 r_{t-1} + \phi_2 r_{t-2}
+ \varepsilon_t$$
The restricted model embodies the martingale -- returns are unpredictable
and equal to a constant mean. The unrestricted model allows the two most
recent lagged returns to enter as predictors. The martingale hypothesis
is equivalent to $H_0: \phi_1 = \phi_2 = 0$, which is tested using a
\textbf{Wald test} with heteroscedasticity-robust standard errors
(vcovHC). Robust standard errors are essential here because financial
returns typically exhibit time-varying volatility, which would otherwise
invalidate standard inference. A p-value below 0.05 constitutes evidence
against the martingale -- i.e. past returns do help predict future returns.

### Results

```{r martingale-results}
martingale_results <- data.frame(
  Ticker     = c("BTC-USD","AMD","KWEB","GLD","CCJ","NEE","MELI","CRSP",
                 "SE","SOL-USD","PLTR","RKLB","ACHR","IONQ","HOOD","NU"),
  F_stat     = c(0.6637,3.3106,0.5600,1.5059,0.2375,1.2383,0.2615,0.3889,
                 0.7912,0.8096,2.8011,0.0883,0.8772,0.2564,0.8593,0.6280),
  p_value    = c(0.5150,0.0367,0.5713,0.2220,0.7886,0.2901,0.7699,0.6779,
                 0.4535,0.4452,0.0612,0.9155,0.4163,0.7739,0.4238,0.5339),
  Martingale = c("Yes","No","Yes","Yes","Yes","Yes","Yes","Yes",
                 "Yes","Yes","Yes","Yes","Yes","Yes","Yes","Yes")
)

kable(martingale_results, booktabs = TRUE,
      caption = "\\label{tab:martingale}Wald Test -- Martingale Hypothesis",
      col.names = c("Ticker","F-statistic","p-value","Martingale")) %>%
  kable_styling(latex_options = c("hold_position","striped"), font_size = 9) %>%
  footnote(general = "Martingale = Yes if p-value $>$ 0.05. Robust standard
           errors (vcovHC) used throughout.", escape = FALSE)
```

The results in Table \ref{tab:martingale} are broadly supportive of the
martingale hypothesis. \textbf{15 out of 16 assets} fail to reject $H_0$
at the 5\% significance level -- knowing the last two days of returns
provides no statistically useful information about tomorrow's return,
precisely what weak-form efficiency predicts.

The sole exception is \textbf{AMD} (F = 3.31, p = 0.037), which rejects
the martingale at the 5\% level. This is somewhat surprising given that
AMD is one of the most liquid and heavily covered stocks in the basket.
One possible explanation is that over a longer sample horizon (2015--2024),
AMD's returns reflect distinct structural regimes -- the company's near-
bankruptcy around 2015, its subsequent turnaround, and the AI-driven rally
from 2023 onwards -- which together generate lagged return patterns that
a linear model can partially detect.

\textbf{PLTR} also warrants attention: with p = 0.061 it sits just
outside the 5\% threshold, suggesting borderline predictability consistent
with its complex ARMA structure identified in the previous section.

Overall, the martingale results confirm that \textbf{daily returns in
this basket are largely unpredictable from their own history}, with AMD
as the most notable exception over the extended sample period.

## RW1: Cowles--Jones and Runs Tests

### Theoretical Framework

The ARMA and martingale tests in the previous section are parametric, 
they assume a linear structure and are sensitive to the distributional
properties of returns. The \textbf{Random Walk 1 (RW1)} hypothesis
imposes a stricter requirement: not only should returns be uncorrelated,
but their increments should be \textit{independently and identically
distributed} (i.i.d.). Testing RW1 therefore calls for nonparametric
methods that are robust to distributional assumptions.

We employ two complementary tests. Both are based on the \textit{sign}
of daily returns -- whether a return is positive or negative -- rather
than its magnitude, making them naturally robust to outliers and
non-normality.

\textbf{The Cowles--Jones Test} (Cowles and Jones, 1937) compares the
frequency of \textit{sequences} -- consecutive returns of the same sign
-- against the frequency of \textit{reversals} -- consecutive returns
of opposite sign. Define the indicator:
$$I_t = \begin{cases} 1 & \text{if } r_t > 0 \\ 0 & \text{if } r_t
\leq 0 \end{cases}$$
A sequence occurs when $I_t = I_{t+1}$, and a reversal when
$I_t \neq I_{t+1}$. The Cowles--Jones ratio is defined as:
$$\widehat{CJ} = \frac{N_s}{N_r}$$
where $N_s$ and $N_r$ are the number of sequences and reversals
respectively. Under RW1 with estimated probability of a positive return
$\hat{\pi}$, the expected ratio equals:
$$E[\widehat{CJ}] = \frac{\pi_s}{1 - \pi_s}, \qquad \pi_s = \hat{\pi}^2
+ (1-\hat{\pi})^2$$
and $\widehat{CJ}$ is asymptotically normally distributed, allowing
construction of a standard $z$-test. A ratio significantly different
from its expected value indicates that the sign of returns is
predictable -- either trending (CJ $> E[CJ]$) or mean-reverting
(CJ $< E[CJ]$).

\textbf{The Runs Test} (Mood, 1940) counts the total number of
\textit{runs} $N_{runs}$ in the return series, where a run is an
uninterrupted sequence of returns with the same sign. Under RW1, the
expected number of runs and its variance are:
$$E[N_{runs}] = 2T\hat{\pi}(1-\hat{\pi}) + \hat{\pi}^2 +
(1-\hat{\pi})^2$$
$$Var[N_{runs}] = 4T\hat{\pi}(1-\hat{\pi})\left[1 - 3\hat{\pi}
(1-\hat{\pi})\right]$$
The test statistic, with continuity correction (Wallis and Roberts,
1956), is:
$$z = \frac{N_{runs} + \frac{1}{2} - 2T\hat{\pi}(1-\hat{\pi})}
{\sqrt{Var[N_{runs}]}} \overset{a}{\sim} \mathcal{N}(0,1)$$
Too few runs indicate positive autocorrelation (trending behaviour);
too many runs indicate negative autocorrelation (mean reversion).
Both tests are applied to all 16 assets and the null of RW1 is
rejected at the 5\% significance level.

### Results

```{r rw1-results}
results_rw1 <- data.frame(
  Ticker    = c("BTC-USD","AMD","KWEB","GLD","CCJ","NEE","MELI","CRSP",
                "SE","SOL-USD","PLTR","RKLB","ACHR","IONQ","HOOD","NU"),
  CJ_ratio  = c(0.8790,0.9744,0.9141,0.9514,0.9901,0.9514,1.0469,0.9438,
                1.0809,0.9597,1.0809,1.0060,1.0638,1.0304,1.1078,0.9819),
  CJ_pvalue = c(0.0208,0.6773,0.1325,0.4127,0.8722,0.4071,0.5118,0.2897,
                0.2022,0.4417,0.2011,0.9588,0.5821,0.6410,0.1181,0.8013),
  CJ_RW1    = c("Reject","Fail to Reject","Fail to Reject","Fail to Reject",
                "Fail to Reject","Fail to Reject","Fail to Reject","Fail to Reject",
                "Fail to Reject","Fail to Reject","Fail to Reject","Fail to Reject",
                "Fail to Reject","Fail to Reject","Fail to Reject","Fail to Reject"),
  RT_pvalue = c(0.0129,0.6504,0.1070,0.3830,0.8468,0.3773,0.5411,0.2596,
                0.2323,0.4171,0.2312,0.9840,0.6104,0.6687,0.1472,0.7718),
  RT_RW1    = c("Reject","Fail to Reject","Fail to Reject","Fail to Reject",
                "Fail to Reject","Fail to Reject","Fail to Reject","Fail to Reject",
                "Fail to Reject","Fail to Reject","Fail to Reject","Fail to Reject",
                "Fail to Reject","Fail to Reject","Fail to Reject","Fail to Reject")
)

kable(results_rw1, booktabs = TRUE,
      caption = "\\label{tab:rw1}RW1 Tests: Cowles--Jones and Runs Test",
      col.names = c("Ticker","CJ Ratio","CJ p-value","CJ Decision",
                    "RT p-value","RT Decision")) %>%
  kable_styling(latex_options = c("hold_position","striped","scale_down"),
                font_size = 9) %>%
  footnote(general = "RW1 rejected at 5\\\\% significance level.",
           escape = FALSE)
```

Table \ref{tab:rw1} presents the results of both tests across all 16
assets. The overall picture is one of broad consistency with the RW1
hypothesis -- \textbf{15 out of 16 assets fail to reject RW1} under
both the Cowles--Jones and runs tests simultaneously.

The sole and striking exception is \textbf{BTC-USD}, which rejects RW1
under both tests (CJ p = 0.021, RT p = 0.013). The CJ ratio of 0.879
is below 1, indicating that \textit{reversals occur more frequently
than sequences} in Bitcoin's daily returns -- positive days tend to be
followed by negative days more often than randomness would predict.
This mean-reverting pattern is consistent with the well-documented
overreaction behaviour in cryptocurrency markets, where retail-driven
sentiment swings cause prices to overshoot and then correct. It is also
noteworthy that this result stands in contrast to the martingale test,
where BTC failed to reject $H_0$ -- a reminder that the two tests
capture different aspects of predictability, and that nonparametric
sign-based tests can detect patterns that linear regression-based tests
miss.

Taken together, the RW1 results reinforce the conclusion from the ARMA
and martingale analysis: \textbf{returns in this basket are largely
consistent with the random walk}, with Bitcoin as the most persistent
anomaly and AMD and PLTR flagged by complementary tests as warranting
closer attention.

## RW2: Market Anomalies

### Theoretical Framework

The \textbf{Random Walk 2 (RW2)} hypothesis relaxes RW1 by allowing
return increments to be non-identically distributed -- accommodating,
for instance, time-varying volatility -- while maintaining the
assumption of \textit{independence}. Testing RW2 therefore shifts focus
from autocorrelation to a different type of predictability: systematic
calendar-based patterns and cross-sectional return regularities known
as \textbf{market anomalies}. If returns are genuinely independent,
no such patterns should exist. We test three anomalies: the
day-of-the-week effect, the January effect, and the winner--loser
effect.

\textbf{The Day-of-the-Week Effect} posits that returns differ
systematically across weekdays -- most notably the so-called
\textit{weekend effect}, whereby Monday returns tend to be lower than
those of other days. If true, this would imply a predictable seasonal
pattern in returns, directly violating independence. We test this using
the same Wald test framework as the martingale test: a restricted model
$r_t = c + \varepsilon_t$ is compared against an unrestricted model
that includes weekday dummies:
$$r_t = c + \sum_{d=2}^{5} \beta_d \mathbf{1}[\text{weekday}_t = d]
+ \varepsilon_t$$
The null hypothesis $H_0: \beta_2 = \cdots = \beta_5 = 0$ is tested
with a Wald test using heteroscedasticity-robust standard errors. A
rejection would indicate that at least one weekday generates
systematically different returns.

\textbf{The January Effect} refers to the empirical regularity,
first documented in the 1970s, whereby stock returns tend to be
abnormally high in January -- particularly for small-cap stocks.
The standard explanation invokes tax-loss selling in December followed
by reinvestment in January. We test this with the same Wald framework,
comparing returns in January against returns in all other months:
$$r_t = c + \beta \mathbf{1}[\text{month}_t = \text{January}] +
\varepsilon_t$$
Under the null $H_0: \beta = 0$, January returns are no different
from the rest of the year.

\textbf{The Winner--Loser Effect}, documented by De Bondt and Thaler
(1985), tests for investor overreaction at a longer horizon. The
hypothesis is that investors systematically overreact to news,
pushing past winners too high and past losers too low. Over a
subsequent period, prices should correct: losers should outperform
and winners should underperform. Following their methodology, stocks
are ranked by cumulative excess return over a \textbf{formation
period} of 12 quarters (3 years), and winner and loser portfolios of
10 stocks each are constructed. Performance is then tracked over the
following \textbf{12-quarter test period} by computing the Average
Cumulative Abnormal Return (ACAR) relative to the S\&P 500 as the
market benchmark. The overreaction hypothesis predicts:
$$ACAR^L_t - ACAR^W_t > 0 \quad \forall t$$
i.e. the loser portfolio should outperform the winner portfolio
throughout the entire test window. To ensure sufficient cross-sectional
variation, the asset universe is expanded to \textbf{66 tickers}
spanning finance, healthcare, energy, consumer staples, industrials,
real estate, materials, utilities, and the original 16 assets, covering
a sample period from \textbf{January 2015 to December 2024}. This
broader universe captures multiple market cycles and a wide range of
industries, making the test more robust and comparable in spirit to
the original De Bondt and Thaler study.

### Day-of-the-Week Effect

```{r dow-results}
dow_results <- data.frame(
  Ticker  = c("BTC-USD","AMD","KWEB","GLD","CCJ","NEE","MELI","CRSP",
               "SE","SOL-USD","PLTR","RKLB","ACHR","IONQ","HOOD","NU"),
  F_stat  = c(0.9387,1.3669,1.5851,0.5050,0.9108,0.9095,0.5203,1.3366,
               1.0173,0.6100,0.3815,0.8112,0.2103,0.6724,0.7683,0.1652),
  p_value = c(0.4658,0.2429,0.1755,0.7321,0.4566,0.4574,0.7208,0.2540,
               0.3970,0.7226,0.8219,0.5181,0.9328,0.6112,0.5460,0.9560),
  DOW     = rep("No Effect", 16)
)

kable(dow_results, booktabs = TRUE,
      caption = "\\label{tab:dow}Wald Test -- Day-of-the-Week Effect",
      col.names = c("Ticker","F-statistic","p-value","Result")) %>%
  kable_styling(latex_options = c("hold_position","striped","scale_down"),
                font_size = 9) %>%
  footnote(general = "Effect present if p-value $<$ 0.05. Robust standard
           errors (vcovHC) used throughout.", escape = FALSE)
```

Table \ref{tab:dow} shows no evidence of a day-of-the-week effect
across any of the 16 assets. All F-statistics are small and all
p-values comfortably exceed the 5\% threshold -- several are above
0.80. This result holds uniformly across very different asset types:
cryptocurrencies that trade on weekends, speculative growth stocks,
defensive commodities, and emerging market equities all fail to show
any systematic weekday pattern. The weekend effect, one of the most
cited anomalies in the early market efficiency literature, finds no
support in this modern, diversified basket.

### January Effect

```{r jan-results}
jan_results <- data.frame(
  Ticker     = c("BTC-USD","AMD","KWEB","GLD","CCJ","NEE","MELI","CRSP",
                 "SE","SOL-USD","PLTR","RKLB","ACHR","IONQ","HOOD","NU"),
  Mean_Jan   = c(-0.1193,0.0890,0.1630,0.1396,0.0565,-0.0016,0.4388,
                  0.3000,0.0619,1.0765,0.3058,-0.1685,-0.5309,-0.3342,
                 -0.2492,-0.1117),
  Mean_Rest  = c(0.1814,0.1577,-0.0121,0.0201,0.0493,0.0540,0.0754,
                 0.0274,0.1081,0.2471,0.1873,0.1189,0.0465,0.1813,
                 0.0309,0.0140),
  p_value    = c(0.2670,0.8184,0.3186,0.0536,0.9743,0.6626,0.0934,
                 0.4675,0.8847,0.2699,0.8195,0.4905,0.2866,0.2997,
                 0.5881,0.8055),
  January    = rep("No Effect", 16)
)

kable(jan_results, booktabs = TRUE,
      caption = "\\label{tab:jan}Wald Test -- January Effect",
      col.names = c("Ticker","Mean Jan (\\%)","Mean Rest (\\%)",
                    "p-value","Result")) %>%
  kable_styling(latex_options = c("hold_position","striped","scale_down"),
                font_size = 9) %>%
  footnote(general = "Mean returns expressed as daily averages in percentage
           terms. Effect present if p-value $<$ 0.05.", escape = FALSE)
```

Table \ref{tab:jan} finds no statistically significant January effect
for any asset in the basket. No p-value falls below the 5\% threshold,
and the direction of the January premium is far from consistent: while
SOL, MELI, CRSP and PLTR show higher average returns in January,
assets like BTC, ACHR, IONQ and HOOD actually perform \textit{worse}
in January than in the rest of the year. Two assets approach
borderline significance -- \textbf{GLD} (p = 0.054) and \textbf{MELI}
(p = 0.093) -- but neither clears the conventional threshold. The
absence of a directional pattern combined with the lack of statistical
significance provides no support for the January effect in this
sample, suggesting that any seasonal regularities documented in earlier
literature have either disappeared or never applied to this type of
asset universe.

### Winner-Loser Effect
```{r winner-loser-plot, fig.cap="Winner--Loser Effect: ACAR over 12-quarter Test Window", out.width="80%"}
plot_data <- data.frame(
  time_step = rep(0:12, 2),
  ACAR      = c(0.00, -0.08, -0.09, -0.02, -0.16, 0.08, 0.01, -0.05,
                0.03, -0.25, -0.20, -0.29, -0.10,
                0.00, -0.07, -0.09, -0.03, -0.02, -0.01, 0.00, -0.01,
                -0.12, -0.15, -0.06, -0.07, 0.04),
  category  = rep(c("Loser","Winner"), each = 13)
)

ggplot(plot_data, aes(x = time_step, y = ACAR, color = category)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 2) +
  scale_color_manual(values = c("Winner" = "royalblue",
                                "Loser"  = "firebrick")) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.5) +
  scale_x_continuous(breaks = 0:12) +
  labs(
    x     = "Quarters After Portfolio Formation",
    y     = "Average Cumulative Abnormal Return (ACAR)",
    color = "Portfolio"
  ) +
  theme_minimal(base_size = 11)
```

Figure \ref{fig:winner-loser-plot} presents results that are mixed and
do \textbf{not cleanly support the overreaction hypothesis} of De Bondt
and Thaler (1985). Both the winner and loser portfolios generate
negative ACARs for most of the test window, suggesting that on average
neither portfolio outperforms the S\&P 500 benchmark over the
2015--2024 sample. The loser portfolio is substantially more volatile,
swinging between brief recoveries and sharp deteriorations, and ends
the test window at approximately $-0.10$ -- below the winner portfolio
which finishes marginally above zero at $+0.04$.

Rather than the clean loser-outperforms-winner pattern predicted by the
overreaction hypothesis, what emerges is a picture of \textbf{persistent
underperformance by past losers} and relative stability for past winners.
This is consistent with a market environment where poor past performers
continue to struggle -- a pattern more aligned with momentum than with
mean reversion. The high volatility of the loser ACAR also suggests
that the loser portfolio is disproportionately composed of distressed or
high-risk assets whose fortunes are highly sensitive to market
conditions, rather than temporarily undervalued stocks poised for
recovery as the overreaction story would predict.

Overall, the winner--loser analysis does not support the overreaction
hypothesis in this sample. If anything, the results lean towards
\textbf{momentum persistence}, where past winners hold their ground
and past losers continue to underperform relative to the market.

## RW3: Return Predictability under the Random Walk 3 Framework

### Theoretical Framework

The \textbf{Random Walk 3 (RW3)} hypothesis is the weakest of the three:
it only requires that return increments are \textit{uncorrelated}, without
imposing independence or identical distributions. Formally, for any
non-zero lag $k$:
$$Cov(r_t, r_{t-k}) = 0 \quad \forall k \neq 0$$
This makes RW3 the most empirically relevant assumption for financial
returns, as it accommodates the time-varying volatility and fat tails
widely observed in practice. The natural test for RW3 is therefore
whether any \textit{linear} autocorrelation remains in the return series
-- precisely what the ARMA analysis and the Ljung--Box test in Section
2.3 were designed to detect. We therefore draw directly on those results
here rather than repeating the analysis.

### Results

The evidence from Section 2.3 already provides a direct test of RW3.
Recall that the Ljung--Box test on the ARMA residuals failed to reject
white noise for 14 out of 16 assets, with only \textbf{MELI}
(p = 0.004) and \textbf{PLTR} (p = 0.037) showing significant residual
autocorrelation. This means that once the modest ARMA structure is
removed, virtually all linear dependence is accounted for -- consistent
with RW3.

To complement this, we apply the Ljung--Box test directly to the
\textbf{raw log-returns} (without any ARMA filtering), which provides
a cleaner and more direct test of the RW3 null:

```{r rw3-results}
# Ljung-Box on raw returns + ACF at lag 1
library(tidyquant)
library(tidyverse)
library(forecast)
library(knitr)
library(kableExtra)

tickers <- c(
  "BTC-USD", "HOOD", "AMD",  "PLTR", "RKLB",
  "IONQ",    "ACHR", "NU",   "SE",   "KWEB",
  "GLD",     "CCJ",  "CRSP", "NEE",  "MELI", "SOL-USD"
)

prices <- tq_get(tickers, from = "2015-01-01", to = "2024-12-31")

returns <- prices %>%
  group_by(symbol) %>%
  arrange(date) %>%
  transmute(date, rtn = log(adjusted) - log(lag(adjusted))) %>%
  na.omit() %>%
  ungroup()

rw3_results <- map_dfr(unique(returns$symbol), function(tk) {
  rtn <- returns %>% filter(symbol == tk) %>% pull(rtn)
  lb  <- Box.test(rtn, lag = 10, type = "Ljung-Box")
  data.frame(
    Ticker   = tk,
    ACF_lag1 = round(acf(rtn, lag.max = 1, plot = FALSE)$acf[2], 4),
    LB_stat  = round(lb$statistic, 4),
    p_value  = round(lb$p.value, 4),
    RW3      = ifelse(lb$p.value > 0.05, "Fail to Reject", "Reject")
  )
})

kable(rw3_results, booktabs = TRUE,
      caption = "\\label{tab:rw3}Ljung--Box Test on Raw Log-Returns (RW3)",
      col.names = c("Ticker","ACF Lag 1","LB Statistic",
                    "p-value","RW3 Decision")) %>%
  kable_styling(latex_options = c("hold_position","striped","scale_down"),
                font_size = 9) %>%
  footnote(general = "Ljung--Box test at lag 10 applied to raw log-returns.
           RW3 rejected at 5\\\\% significance level.", escape = FALSE)
```

Table \ref{tab:rw3} presents a more mixed picture than earlier tests.
\textbf{8 out of 16 assets reject RW3} -- BTC, AMD, NEE, MELI, CRSP,
SOL, PLTR, and NU -- while the remaining 8 fail to reject. This is a
stronger result against the random walk than the martingale and RW1
tests, where only 1--2 assets showed significant predictability.
\textbf{PLTR} is the most consistently predictable asset across all
tests, with the highest lag-1 ACF (0.087) and LB statistic (39.57,
p $\approx$ 0). \textbf{NEE} is the most surprising rejection, with
an LB statistic of 68.03 despite showing no anomalies in earlier
tests. That said, all lag-1 ACF coefficients are small in magnitude --
rarely exceeding 0.05 -- suggesting that while autocorrelation is
statistically detectable, it is unlikely to be economically exploitable
after transaction costs.